---
title: "Calorimetry Lab Report"
author: "Andrés Peña"
date: "3/2/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(readr)
library(ggplot2)
library(magrittr)
library(dplyr)
library(lubridate)
library(latex2exp)
knitr::opts_chunk$set(echo = FALSE)

benzoic_acid <- read_csv("benzoic acid.csv")
naphthalene <- read_csv("naphthalene.csv")
gummy_bear <- read_csv("gummy bear.csv")
```

*Heat capacity of the system*
\newline

```{r}
m_wire <- 0.0185
m_wire_left <- 0.0079
m_wire_combusted <- m_wire - m_wire_left
m_wire_pellet <- 1.0126
m_pellet <- m_wire_pellet - m_wire
molar_mass_BA <- 122

time_combustion = ms("43:30")
```

Mass of wire = `r m_wire` g

Mass of wire left = `r m_wire_left` g

Mass of wire combusted = `r m_wire_combusted` g

Mass of wire + pellet = `r m_wire_pellet` g

Mass of pellet = `r m_pellet` g

```{r echo=FALSE}
delta_h_BA <- m_pellet * -26435.8
delta_u_wire <- m_wire_combusted * -5858

mol_BA <- m_pellet / molar_mass_BA
mol_gas_net <- 2.5
delta_n <- mol_gas_net * mol_BA

gas_constant <- 8.3145
temp <- 298.2
delta_u_BA <- delta_h_BA - gas_constant * temp * delta_n
```

$\Delta H_c (BA) = 0.9941g\ *\ -26435.8\ J/g =$ `r delta_h_BA` J

$\Delta U (wire) = 0.0185\ g\ * -5858\ J/g =$ `r delta_u_wire` J

$C_6 H_5 COOH(s) + 7.5O_2 (g) \rightarrow 7CO_2(g) + 3H_2 O(g)$

Moles of benzoic acid = `r mol_BA`

Moles of net gas produced per mole of benzoic acid = 2.5

$\Delta n$ = 2.5 * Moles of benzoic acid = `r delta_n`

$\Delta U(BA) = \Delta H(BA) - RT \Delta\ n$ = `r delta_u_BA` J


```{r, echo = FALSE}

max_row <- benzoic_acid %>%
  arrange(desc(temperature)) %>%
  head(1)
max <- max_row[['time']]

benzoic_acid_cooling <- benzoic_acid %>%
  filter(time > max)

cooling_m <- lm(temperature ~ time, data = benzoic_acid_cooling)
benzoic_acid <- benzoic_acid %>%
  mutate(
    pred_temp = predict(cooling_m, benzoic_acid)
  )

combustion <- benzoic_acid %>%
  filter(time == time_combustion)
max_t <- combustion[['pred_temp']]
min_t <- combustion[['temperature']]
delta_t <- max_t - min_t

benzoic_acid_plot <- ggplot(data = benzoic_acid) +
  geom_point(aes(x = time, y = temperature)) +
  geom_point(aes(x = time, y = pred_temp), color = "blue") +
  geom_segment(
    aes(x = time_combustion,
        y = min_t,
        xend = time_combustion,
        yend = max_t),
    arrow = arrow()
  ) +
  geom_text(aes(x = time_combustion - ms("1:00"),
                y = min_t + (max_t - min_t) / 2),
            label = "Delta T") +
  labs(
    title = "Benzoic Acid Combustion",
    x = "Time",
    y = "Temperature (C)"
  )
benzoic_acid_plot

c_v <- -(delta_h_BA - gas_constant * temp * delta_n + delta_u_wire) / delta_t
```

$C_v = -\frac{\Delta H(BA)\ -\ RT\Delta n\ +\ \Delta U (Wire)}{\Delta T}$

$C_v=$ `r c_v` J/K

\newpage

*Naphthalene*
\newline

```{r}
m_wire <- 0.0153
m_wire_left <- 0.0136
m_wire_combusted <- m_wire - m_wire_left
m_wire_pellet <- 0.5748
m_pellet <- m_wire_pellet - m_wire
ratio_mol_gas <- 1.5
molar_mass_naph <- 128.171

time_combustion = hms("1:31:40")
```

Mass of wire = `r m_wire` g

Mass of wire left = `r m_wire_left` g

Mass of wire combusted = `r m_wire_combusted` g

Mass of wire + pellet = `r m_wire_pellet` g

Mass of pellet = `r m_pellet` g

```{r echo=FALSE}
delta_h_ <- m_pellet * -26435.8
delta_u_wire <- m_wire_combusted * -5858

mol_naph <- m_pellet / molar_mass_naph
mol_gas_net <- 2
delta_n <- mol_gas_net * mol_BA
```

$\Delta U (wire) =$ `r m_wire_combusted` g * -5858 J/g = `r delta_u_wire` J

$C_{10} H_8(s) + 12O_2 (g) \rightarrow 10CO_2(g) + 4H_2 O(g)$

Moles of naphthalene = `r mol_BA`

Moles of net gas produced per mole of naphthalene = 2

$\Delta n$ = 2 * Moles of naphthalene = `r delta_n`

```{r, echo = FALSE}

max_row <- naphthalene %>%
  arrange(desc(temperature)) %>%
  head(1)
max <- max_row[['time']]

naphthalene_cooling <- naphthalene %>%
  filter(time > max)

cooling_m <- lm(temperature ~ time, data = naphthalene_cooling)
naphthalene <- naphthalene %>%
  mutate(
    pred_temp = predict(cooling_m, naphthalene)
  )

combustion <- naphthalene %>%
  filter(time == time_combustion)
max_t <- combustion[['pred_temp']]
min_t <- combustion[['temperature']]
delta_t <- max_t - min_t

naphthalene_plot <- ggplot(data = naphthalene) +
  geom_point(aes(x = time, y = temperature)) +
  geom_point(aes(x = time, y = pred_temp), color = "blue") +
  geom_segment(
    aes(x = time_combustion,
        y = min_t,
        xend = time_combustion,
        yend = max_t),
    arrow = arrow()
  ) +
  geom_text(aes(x = time_combustion - ms("0:30"),
                y = min_t + (max_t - min_t) / 2),
            label = "Delta T") +
  labs(
    title = "Naphthalene Combustion",
    x = "Time",
    y = "Temperature (C)"
  )
naphthalene_plot

delta_h_naph = - delta_t * c_v + gas_constant * temp * delta_n - delta_u_wire

delta_h_naph_molar = delta_h_naph / mol_naph
```

$\Delta H(N) = -\Delta TC_v + RT\Delta n - \Delta U (wire)$

$\Delta H(N)=$ `r delta_h_naph` J

$\Delta \tilde H (N) = \frac {\Delta H(N)}{moles\ naphthalene}$

$\Delta \tilde H (N)=$ `r delta_h_naph_molar` J/mol

\newpage

*Gummy Bear*
\newline

```{r}
m_wire <- 0.0162
m_wire_left <- 0.0080
m_wire_combusted <- m_wire - m_wire_left
m_wire_pellet <- 2.2202
m_pellet <- m_wire_pellet - m_wire

time_combustion = hms("1:59:20")
```

Mass of wire = `r m_wire` g

Mass of wire left = `r m_wire_left` g

Mass of wire combusted = `r m_wire_combusted` g

Mass of wire + gummy bear = `r m_wire_pellet` g

Mass of gummy bear = `r m_pellet` g

```{r echo=FALSE}
delta_h_ <- m_pellet * -26435.8
delta_u_wire <- m_wire_combusted * -5858
```

$\Delta U (wire) =$ `r m_wire_combusted` g * -5858 J/g = `r delta_u_wire` J

```{r, echo = FALSE}

max_row <- gummy_bear %>%
  arrange(desc(temperature)) %>%
  head(1)
max <- max_row[['time']]

gummy_bear_cooling <- gummy_bear %>%
  filter(time > max)

cooling_m <- lm(temperature ~ time, data = gummy_bear_cooling)
gummy_bear <- gummy_bear %>%
  mutate(
    pred_temp = predict(cooling_m, gummy_bear)
  )

combustion <- gummy_bear %>%
  filter(time == time_combustion)
max_t <- combustion[['pred_temp']]
min_t <- combustion[['temperature']]
delta_t <- max_t - min_t

gummy_bear_plot <- ggplot(data = gummy_bear) +
  geom_point(aes(x = time, y = temperature)) +
  geom_point(aes(x = time, y = pred_temp), color = "blue") +
  geom_segment(
    aes(x = time_combustion,
        y = min_t,
        xend = time_combustion,
        yend = max_t),
    arrow = arrow()
  ) +
  geom_text(aes(x = time_combustion - ms("0:30"),
                y = min_t + (max_t - min_t) / 2),
            label = "Delta T") +
  labs(
    title = "Gummy Bear Combustion",
    x = "Time",
    y = "Temperature (C)"
  )
gummy_bear_plot

delta_u <- - delta_t * c_v - delta_u_wire

```

$\Delta U(bear) = -\Delta TC_v - \Delta U (wire)$

$\Delta U(bear)=$ `r delta_u` J